HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-07-25 scnn.tex
updated refs

\begin{keyword}
Spherical Convolutional Neural Network \sep SCNN \sep HealpixNet \sep Graph CNN \sep \nati{@Tomek: could you put some cosmo keywords?}
\end{keyword}

@@

\section{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
    \item (NP) Figures  A.15, 6, 5
    \item (TK) check all the colored stuff left for you
    \item (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook.
    \item (TK) update the README for data zip - file from NP
    \item (NP) Finalise readme on Github code
    \item (Waiting for Zenodo) (NP) Finalize the data on zenodo
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item NP move the repo to HealpixNet
    \item (TK) make the bibliography display only a few name
	\item check if papers which appear as arxiv were not published in a conference or journal
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (TK) possibly improve plot of pooling
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    \item (first version done) conclusions
    \item Should the heat example go to appendix?
    \item TK read the heat section again
    \item TK read the result section
    \item TK check the appendix
    \item TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item TK read everything
	\item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings.
	\item uniformize text size in figures
\end{itemize}

@@

\todo{@Tomek: can you explain why we care about parts of the sphere? (that is an advantage of the graph model over the spherical harmonics)}
\todo{@Tomek: How is the data acquired? Do telescopes and satellites take measurements according to the HEALPix sampling? (it is also an advantage of our method to be agnostic to the sampling, which might be useful to researchers working with missing data or different models)}
\nati{@Tomek: I think what Michael has in mind is that the graph approach is very versatile and extendable. It can be used with other samplings (even irregular ones) or only part of the sphere. This flexibility might be an advantage to mention. In particular, compared to~\citet{cohen2018spherical} it allows us to do only do the required computation when only a part of the sphere is considered.}

So far these algorithms have mostly been demonstrated on Euclidean domains, such as flat images.
As such, perhaps the most straightforward approach to the analysis of spherical data is to divide it into small chunks and project those on flat 2D surfaces.
While this mostly works, as demonstrated \todo{in ??? for example}, \nati{@Tomek, do you have a citation for that? Otherwise, we need to reformulate this.} \nati{Cite the paper of Janis?} it is more natural and accurate to work directly on the sphere.
The main challenge in designing a CNN on the sphere is to define a convolution which is suitable to this irregularly sampled domain.

@@

For all the
experiments, we work with centered Gaussian noise. \todo{Tomek: could you put a
justification.} To improve the
stability of the spherical CNN, we slowly increase the amount of noise added to
the training data over the iterations. Furthermore in order to avoid
overfitting, the noise is randomly produced during the training process. We trained our
model between $50$ and $290$ epochs depending on the complexity and the size of
the problem. $20\%$ of the training set that we call "validation set" was left out of the training process and simply used to assess the global performance and overfitting of
the network.
Training took less than a day for each model using an NVIDIA Tesla K20.

@@

\subsection{Filter visualization}
One common visualization for CNN is to observe the shape of the learned filters.
Since our construction leads to almost (up to the sampling irregularities)
spherical filters, we plot in \figref{learned_filter} both the section (on the
sphere) and a gnomonic projection on a plane of a random selection of learned
filters. The filters are selected from the last layer of the network
corresponding the experiment with order $o=2$ for a relative noise level of
$2$. Nevertheless, all networks presented similar gnomonic and section pattern, at least from a
human perspective. The gnomonic projection clearly exibits the sampling issue
of HEALPix. For a detail, on how the convolution kernels are plotted, please refer to Appendix~\ref{sec:filter_visualization}.
\todo{Tomek: can you say something about the filters.}

@@

As shown in this contribution, modeling the discretized sphere, a 2D manifold in $\R^3$, by a graph enables an efficient convolution operation on any pixelization. Hence, it relaxes the iso-latitude constraint on sampling schemes who aim to enable fast convolutions. For applications which can tolerate small imprecisions in rotation invariance --- the fact that the action of a graph filter slightly depends on the location --- this relaxation might enable researchers to consider sampling sets with different properties.

>>>>

\begin{keyword}
Spherical Convolutional Neural Network \sep
HealpixNet \sep
Graph CNN \sep
Cosmological data analysis \sep
Mass mapping
\end{keyword}

@@

\section{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
    \item (NP) Figures  A.15, 6, 5
    \item (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook.
    \item (TK) update the README for data zip - file from NP
    \item (NP) Finalise readme on Github code
    \item (Waiting for Zenodo) (NP) Finalize the data on zenodo
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item NP move the repo to HealpixNet
	\item check if papers which appear as arxiv were not published in a conference or journal
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    \item (first version done) conclusions
    \item Should the heat example go to appendix?
    \item TK read the heat section again
    \item TK read the result section
    \item TK check the appendix
    \item TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item TK read everything
	\item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings.
	\item uniformize text size in figures
\end{itemize}

@@

For cosmological observations it is often the case that the data covers only a part of the sphere.
For ground-based observations this can be due to limited visibility of the sky from a particular telescope location and survey strategy, and for space-based imagers due to masking of the galactic plane area.

% \todo{@Tomek: can you explain why we care about parts of the sphere? (that is an advantage of the graph model over the spherical harmonics)}
% \todo{@Tomek: How is the data acquired? Do telescopes and satellites take measurements according to the HEALPix sampling? (it is also an advantage of our method to be agnostic to the sampling, which might be useful to researchers working with missing data or different models)}
% \nati{@Tomek: I think what Michael has in mind is that the graph approach is very versatile and extendable. It can be used with other samplings (even irregular ones) or only part of the sphere. This flexibility might be an advantage to mention. In particular, compared to~\citet{cohen2018spherical} it allows us to do only do the required computation when only a part of the sphere is considered.}

So far these algorithms have mostly been demonstrated on Euclidean domains, such as flat images.
As such, perhaps the most straightforward approach to the analysis of spherical data is to divide it into small chunks and project those on flat 2D surfaces.
While this mostly works, as demonstrated by \citep{fluri2018deep,gupta2018nongaussianinformation,schmelze2017cosmologicalmodel,gillet2018deeplearning}, it may be more natural for some applications and accurate to work directly on the sphere.
The main challenge in designing a CNN on the sphere is to define a convolution which is suitable to this irregularly sampled domain.

@@

For all the experiments, we work with centered Gaussian noise.
While the noise model is real mass maps often has a slightly different distribution, Gaussian should be a sufficient model to demonstrate the performance of our method.
To improve the stability of the spherical CNN, we slowly increase the amount of noise added to
the training data over the iterations. Furthermore in order to avoid
overfitting, the noise is randomly produced during the training process. We trained our
model between $50$ and $290$ epochs depending on the complexity and the size of
the problem. $20\%$ of the training set that we call "validation set" was left out of the training process and simply used to assess the global performance and overfitting of
the network.
Training took less than a day for each model using an NVIDIA Tesla K20.

@@

\subsection{Filter visualization}
One common visualization for CNN is to observe the shape of the learned filters.
Since our construction leads to almost (up to the sampling irregularities)
spherical filters, we plot in \figref{learned_filter} both the section (on the
sphere) and a gnomonic projection on a plane of a random selection of learned
filters. The filters are selected from the last layer of the network
corresponding the experiment with order $o=2$ for a relative noise level of
$2$. Nevertheless, all networks presented similar gnomonic and section pattern, at least from a
human perspective. The gnomonic projection clearly exibits the sampling issue
of HEALPix.
For a detail, on how the convolution kernels are plotted, please refer to Appendix~\ref{sec:filter_visualization}.
While it is usually difficult to interpret the shape of the filters in the context of the type of data, we can notice that they often have a ``peak'' like structures.
An example of filter interpretation was demonstrated in \citep{Ribli2018learningfrom}.
% \todo{Tomek: can you say something about the filters.}
% \TK{Done }

@@

As shown in this contribution, modelling the discretized sphere, a 2D manifold in $\R^3$, by a graph enables an efficient convolution operation on any pixelization. Hence, it relaxes the iso-latitude constraint on sampling schemes who aim to enable fast convolutions. For applications which can tolerate small imprecisions in rotation invariance --- the fact that the action of a graph filter slightly depends on the location --- this relaxation might enable researchers to consider sampling sets with different properties.
